---
title: "Biological Records Systematic Review - Preliminary Results"
output:
  html_document: default
  html_notebook: default
  word_document: default
---

```{r, echo=F}
knitr::opts_chunk$set(echo=F, message=F, warning=T, cache=F)
```

```{r, load_data, cache = F, results='hide', warning = F}
setwd("~/Documents/Data_Analysis/UCD/systematic_review/")
library(wgutil)
library(Hmisc)
library(captioner)
library(knitr)
library(pander)
library(ggridges)
library(bestglm)
library(car)
library(GGally)
library(tidyverse)

diag <- F # produce exploratory & diagnostics plots & results

# read in data
elig <- read_csv("~/Documents/UCD/PhD_Project/systematic_review/master_eligibility_results.csv")
rev <- read_csv("./data/systematic_review_coded_results.csv")
# remove columns which are used as visual separators with no data
if(any(grepl("X.*", colnames(rev)))) {
  rev <- rev[, -which(grepl("X.*", colnames(rev)))]
}

colnames(rev) <- make.names(colnames(rev))

# tidy some column names
colnames(rev)[which(colnames(rev) ==
                      "X.if.prediction.used..prediction.performance.measure")] <- 
  "predicton.performance.measure"

rev <- rev[which(rev$coding.DONE == T), ] 
rev <- rev[which(rev$qualifies == T), ]
rev <- rev[, -which(colnames(rev) %in% c(
  "link", "inclusion.order", "other.notes", 
  "aspects.not.yet.coded", 
  "eligible.for.meta..or.publication.bias.analysis", 
  "wg.needs.to.read"))]
```

```{r calc_temp_extent}
# calculate number of years covered by study
rev$temp_extent <- as.numeric(rev$end.year) - as.numeric(rev$start.year)
# individually assign values to studies with temporal extent < 1 year
if(diag) rev$title[rev$temp_extent == 0] 
rev$temp_extent[which(
  grepl("An assessment of bumblebee .* land use and floral.*", 
        rev$title))] <- 0.33
```

```{r museum_data_to_whwhwh}
## For the purposes of analysis, if there was a specimen available but the 
# specimen was not used at all for analysis, then, if there are no other richer
# data types, I will consider the data what, where, when-only.  I think this
# best represents the use of museum data.  I do not think this would make sense
# for other data types (e.g. if there is abundance data but the analysis only
# uses it as what, where, when) because that is a choice to strip usually 
# meaningful information.  The physical specimen is likely not meaningful for
# many analyses that use museums as a source of wh, wh, wh data.

for(i in 1:nrow(rev)) {
  if(rev$data.type...museum[i] == T | rev$data.type...photo[i] == T | 
     rev$data.type...audio[i] == T | rev$data.type...video[i] == T) {
    # If a voucher is available
    if(rev$data.type...abundance[i] == F & 
       rev$data.type...sampling.effort.reported[i] == F & 
       rev$data.type...organized.data.collection.scheme[i] == F & 
       rev$data.type...visit.specific.covariates[i] == F & 
       rev$data.type...life.stage[i] == F) {
      # If there are no other rich data types
      if(rev$voucher.of.some.kind.necessary.for.analysis[i] == F) {
        # If the voucer was not used
              rev$data.type...what.where.when.only[i] <- TRUE
      }
    }
  }
} # my, what a very large "if" nest you have, Grandmother.
```

```{r numberingPrep}
# make functions for adding numbered captions to figures 
figs <- captioner(prefix = "Fig.") 
```

# Preliminary structure of Results
This document shows the tables and graphs that I expect to want in the *Results* section of the biological records systematic review.  These are currently made using `r nrow(rev)` articles that I've coded so far.  

## Number of eligible studies
So far, we have evaluated `r sum(!is.na(elig$qualifies))` studies (of `r nrow(elig)`) for eligibility, and judged `r length(which(elig$qualifies == TRUE))` to be eligible for inclusion in the scoping review.

## General Methods

#### Data Type
We coded eleven variables describing data type: what, where, when only; sampling effort reported; abundance; detection / non-detection; organized monitoring scheme; visit-specific covariates; life stage information; museum specimen; photo; audio; video. We treated what, where, when as the baseline data type and considered all other data types supplementary additions to that basic data type.  For statistical analyses, we used only the sampling effort reported, abundance, and detection / non-detection data type variables based on *a-priori* expectations about their expected benefit in enabling inferential and predictive modelling and their expected cost in terms of complexity in collecting, recording, and storing the data.  

## Authors and Data Providing Institutions

**Who choses to use biological records data?**

$H_a$: The first author is more likely to use data held by their own institution than by other institutions.

$H_0$: The number of studies with a lead author from the institution that provided the data is the same as expected by chance.

*Proposed Test*: Permutation test of lead author institutions.  

## Temporal Extent

$H_a$: The mean temporal extent of studies using only what, where, when data is longer than the mean temporal extent of studies using richer data types. 

$H_0$: The mean temporal extent of studies using what, where, when-only data is the same as the mean temporal extent of studies using richer data types. 

*Proposed Test*: Linear regression with natural-log transformed time (in years) as the outcome and data types as predictors.

### Methods

We used linear regression with natural-log transformed time (in years) as the response variable and the following data types as predictors:  sampling effort reported, detection / non-detection, abundance.  Note that these are each individual variables - this is not a single categorical factor level "data type" variable because the variables are not mutually exclusive.  This model assumes an additive effect of each supplementary type of data.  The intercept-only model is data that is just what, where, when only and the model estimates the effect of adding each additional type of supplementary data.  

```{r prepare_temporal_extent_df}
temp_extent <- select(rev, c(title, temp_extent, 
                             data.type...what.where.when.only, 
                             data.type...detection...non.detection, 
                             data.type...abundance,
                             data.type...sampling.effort.reported)) %>%
  gather(key = "data_type", value = "dt_value", 
         data.type...what.where.when.only:data.type...sampling.effort.reported, 
         factor_key = T) %>%
  filter(dt_value == T) %>%
  group_by(data_type)

# make names prettier
temp_extent$data_type <- gsub("data.type...", "", temp_extent$data_type)

temp_extent$data_type <- factor(as.character(temp_extent$data_type), 
                            levels = c("what.where.when.only", 
                                       "sampling.effort.reported",
                                       "detection...non.detection", 
                                       "abundance"), 
                            labels = c("what where when only",
                                       "sampling effort reported", 
                                       "detection / non-detection", 
                                       "abundance"))
```

```{r temp_range_distribution}
if(diag) {
  # distribution of all data
  hist(rev$temp_extent)
  hist(rev$temp_extent[which(rev$temp_extent <= quantile(rev$temp_extent, 
                                                         0.90, na.rm = T))])
  ## data are not balanced
  table(temp_extent$data_type)
  
  ggplot(data = temp_extent, aes(y = temp_extent, x = data_type)) + 
    geom_boxplot() +
    theme_bw() + 
    ggtitle("All data") + 
    theme(axis.text.x = element_text(angle = 330, vjust = 0.3))
  
  ggplot(data = temp_extent, aes(y = temp_extent, x = data_type)) + 
    ylim(c(0, 800)) + 
    geom_boxplot() +
    theme_bw() + 
    ggtitle("Temporal Duration after\ncutting off the 8000 year study") + 
    theme(axis.text.x = element_text(angle = 330, vjust = 0.3))
  
  ggplot(data = temp_extent, aes(x = temp_extent, y = data_type)) + 
    xlim(c(0, 800)) + 
    geom_density_ridges() +
    ggtitle("Temporal Duration after\ncutting off the 8000 year study") + 
    theme_bw() 
  
  ## normality -------------------------------------
  # data very non-normal (above) but log transformation kinda fixes this
  # (reciprocal and sqrt transformations did not help)
  print(ggplot(data = temp_extent, aes(y = temp_extent, x = data_type)) + 
          geom_boxplot(varwidth = T) +
          scale_y_log10() + 
          theme_bw() + 
          ggtitle("log transformation of year\nmakes normality better") + 
          theme(axis.text.x = element_text(angle = 330, vjust = 1, hjust = 0)) 
  )
  
  for(i in unique(temp_extent$data_type)) { 
    # distribution of log(years) by data type
    hist(log(temp_extent$temp_extent)[which(temp_extent$data_type == i)])
  }
  ## normality ok-ish after log trans ---------------
  
  ggplot(data = temp_extent, aes(x = temp_extent, y = data_type)) + 
    scale_x_log10() + 
    # geom_density_ridges(stat = "binline") +
    geom_density_ridges() + 
    theme_bw()

### assess correlation between data type predictors ----------------------
print(ggpairs(rev, columns = c(38, 39, 41, 42)))
}
```

Potential correlations between the following data types: abundance & detection/non-det, sampling effort, organized data collection scheme; sampling effort & det/non-det, organized data collection. **Need to revisit this and decide if I should continue using these *a-priori* variables or whether I should switch b/c of correlation.**

```{r temp_extent_lm_independent_predictors, include = F}
## using original df and each data type as a separate variable 
## fit full model
rev$log_years <- log(rev$temp_extent)
time_lm_full <- lm(log_years ~ 1 + 
                     data.type...what.where.when.only + 
                     data.type...sampling.effort.reported + 
                     data.type...abundance +
                     data.type...detection...non.detection,
                   data = rev) # variables added most to least important
time_resids <- rstandard(time_lm_full)
if(diag) {
  plot(time_lm_full)
  hist(time_resids)
  boxplot(time_resids)
}
time_summary <- summary(time_lm_full)
# time_full_model_anova <- anova(time_lm_full) # not recommended because R guesses a null model
time_summary

## to interpret outcome on original scale:
# for a one-unit increase in x, the predicted value of y 
# changes by (e^beta1 - 1)*100 percent
```

### Results for Time

```{r time_data_type_full_model_sig_test, include = F}
# Test significance of full model
m0 <- lm(log_years ~ 1, data = rev) # fit null model 
m_time_full_anova <- anova(m0, time_lm_full)
m_time_full_anova
```

```{r time_data_type_test_main_effects, include = F}
## test effect of each data type predictor by comparing the full model to a
# model with that predictor removed
# 
# TODO: Does this need multiple comparison correction?

m_no_whwhwh <- lm(log_years ~ 1 + 
                    data.type...sampling.effort.reported + 
                    data.type...abundance +
                    data.type...detection...non.detection,
                  data = rev)
m_no_effort <- lm(log_years ~ 1 + 
                    data.type...what.where.when.only + 
                    data.type...abundance +
                    data.type...detection...non.detection,
                  data = rev)
m_no_abund <- lm(log_years ~ 1 + 
                   data.type...what.where.when.only + 
                   data.type...sampling.effort.reported + 
                   data.type...detection...non.detection,
                 data = rev)
m_no_nonDet <- lm(log_years ~ 1 + 
                    data.type...what.where.when.only + 
                    data.type...sampling.effort.reported + 
                    data.type...abundance,
                  data = rev)
indiv_mods_time <- list(m_no_whwhwh, m_no_effort, m_no_abund, m_no_nonDet)

for(i in 1:length(indiv_mods_time)) { 
  print(anova(indiv_mods_time[[i]], time_lm_full))
}
```

The overall model using supplementary data types to predict the natural log of temporal extent of study (in years) was significant ($F_{`r format(round(m_time_full_anova$Df[2], digits = 2), scientific = F)`, `r format(round(m_time_full_anova$Res.Df[2], digits = 2), scientific = F)`}$ = `r format(round(m_time_full_anova$F[2], digits = 2), scientific = F)`, p = `r format(round(m_time_full_anova[6][[1]][2], digits = 4), scientific = F)`, Adjusted $R^2%$ = `r format(round(time_summary$adj.r.squared, digits = 2), scientific = F)`).  

If the overall model is significant, I will report individual data type results in a format like this: 
 
The expected temporal extent covered by a study that uses only "what, where, when" data is `r round(exp((time_lm_full$coefficients[1] + time_lm_full$coefficients[2])), digits = 1)` years. 

The expected temporal extent of studies is shorter for studies that include sampling effort information than for studies that use only "what, where, when" data (*p* = `r format(round(time_summary$coefficients[2, 4], digits = 3), scientific = F)`).  The effects of additional information on detection / non-detection and abundance were not significant after accounting for the effects of the other data types.  The signs of the effects of sampling effort information and detection / non-detection information were both negative, as expected.  However, the sign of the effect for abundance information was unexpectedly positive.  This may be because abundance is correlated with detection/non-detection and sampling effort.  When I fit a model with only the intercept and the abundance data type, the direction of the effect of abundance data was negative as expected.    

```{r time_extent_results_plot}
## TODO: need to hand-annotate R2 and p-value in plot
temp_extent$sig_different_from_WhWhWh <- temp_extent$data_type %in% 
                                       c("sampling effort reported")
print(ggplot(data = temp_extent, aes(y = temp_extent, x = data_type)) + #, color = sig_different_from_WhWhWh
        geom_boxplot(varwidth = T) +
        scale_y_log10() + 
        xlab("Data Type") + 
        ylab("Temporal Extent of Study (years)") + 
        annotate("text", x = 3, y = 2600, 
                 label = "paste(\"Adjusted \", 
                 italic(R) ^ 2, \" = 0.3\")", 
                parse = TRUE) + 
  annotate("text", x = 3, y = 1200, 
           label = "paste(\"Overall model significance \", 
                 italic(p), \" = 0.012\")", 
           parse = TRUE) +
        theme_bw() + 
        scale_color_discrete(name = "Significantly Different from\nWhat, Where, When-only\nstudies (alpha = 0.1 level)") + 
        theme(axis.text.x = element_text(angle = 330, vjust = 1, hjust = 0)) 
)
```

```{r time_extent_results_plot_for_presentation_slide, eval = F, include = F}
## This code is to be run in the console for making plot to be exported as jpg for slides
## TODO: need to hand-annotate R2 and p-value in plot
t_size = 22 # plot text size
ggplot(data = temp_extent, aes(y = temp_extent, x = data_type)) + #, color = sig_different_from_WhWhWh
  geom_boxplot(varwidth = T) +
  scale_y_log10() + 
  xlab("Data Type") + 
  ylab("Temporal Extent of Study (years)") + 
  # annotate("text", x = 3, y = 140, 
  #          label = "paste(\"Adjusted \", 
  #                italic(R) ^ 2, \" = 0.27\")", 
  #          parse = TRUE, 
  #          size = t_size - 15) + 
  annotate("text", x = 3, y = 1400, 
           label = "paste(\"Overall model significance \", 
                 italic(p), \" = 0.012\")", 
           parse = TRUE, 
           size = t_size - 14) +
  theme_bw() + 
  scale_color_discrete(name = "Significantly Different from\nWhat, Where, When-only\nstudies (alpha = 0.1 level)") + 
  theme(axis.text.x = element_text(angle = 330, vjust = 1, hjust = 0), 
        text = element_text(size = t_size)) 
```


## Study Question Paradigm

**What is the most popular study question paradigm?**

$H_a$: Some broad study question paradigms are more common than others in studies using biological records.

$H_0$: The number of studies pursuing each of the broad study question paradigms is the same.

*Proposed Test* One Poisson regression with study question paradigm as predictor. 

```{r make_study_question_paradigm_tibble}
paradigm_df <- select(rev, title, individual.species.analysis, 
                      community.analysis) %>%
  gather(key = "paradigm", value = "used", individual.species.analysis:
           community.analysis) %>%
  filter(used == TRUE)

table(paradigm_df$paradigm)
```

**What data types are used for each study question paradigm?**
[ descriptive results only ] 

```{r make_study_question_paradigm_by_dataType_tibble}
paradigm_data_df <- select(rev, title, individual.species.analysis, 
                      community.analysis, data.type...what.where.when.only, 
                      data.type...sampling.effort.reported, 
                      data.type...detection...non.detection, 
                      data.type...abundance) %>%
  gather(key = "paradigm", value = "parad_used", individual.species.analysis:
           community.analysis) %>%
  filter(parad_used == TRUE) %>%
  gather(key = "data_type", value = "dat_used", 
         data.type...what.where.when.only:data.type...abundance) %>%
  filter(dat_used == TRUE)

paradigm_data_df$data_type <- gsub("data.type...what.where.when.only", 
                                   "what, where, when only", 
                                   paradigm_data_df$data_type)
paradigm_data_df$data_type <- gsub("data.type...sampling.effort.reported", 
                                   "sampling effort", 
                                   paradigm_data_df$data_type)
paradigm_data_df$data_type <- gsub("data.type...detection...non.detection", 
                                   "detection / non-detection", 
                                   paradigm_data_df$data_type)
paradigm_data_df$data_type <- gsub("data.type...abundance", 
                                   "abundance", 
                                   paradigm_data_df$data_type)
paradigm_data_df$paradigm <- gsub("community.analysis", 
                                   "community", 
                                   paradigm_data_df$paradigm)
paradigm_data_df$paradigm <- gsub("individual.species.analysis", 
                                   "individual species", 
                                   paradigm_data_df$paradigm)

kable(table(paradigm_data_df$data_type, paradigm_data_df$paradigm))
```


## Specific Study Focus

Descriptive stats for studies focusing on policy or alien species, testing ecological theory, and testing methods of data correction.

```{r specific_question_by_paradigm_bar_plot}
qt_df <- select(rev, "title", 
                "species.richness", "diversity", 
                "distribution", "abundance", "phenology") 
names(qt_df)[2:6] <- c("species.richness", "diversity", 
                        "distribution", "abundance", "phenology")

qt_df <- gather(qt_df, key = "study_question", value = "v_2", 
                species.richness, diversity, distribution, abundance, 
                phenology,
                factor_key = T) %>%
  filter(v_2 == TRUE) %>%
  group_by(study_question) %>%
  summarise(count = n()) %>%
  mutate(percent = (count/sum(count)) * 100) %>%
  mutate(paradigm = str_replace_all(study_question, 
                                    c("species.richness|diversity" = 
                                        "community", 
                                      "distribution|abundance|phenology" = 
                                        "individual species"))) 

ggplot(data = qt_df, 
       aes(x = factor(study_question, 
                      levels = c("species.richness", "diversity", 
                                 "distribution", "phenology", "abundance"), 
                      labels = c("species richness", "diversity", 
                                 "distribution", "phenology", "abundance")), 
           y = percent, 
           fill = paradigm)) + 
  geom_bar(stat = "identity") + 
  scale_fill_brewer() + 
  xlab("Study Focus") + 
  ylab("Percent of Studies") + 
  theme_bw()
```

```{r specific_question_by_paradigm_bar_plot}
qt_df <- select(rev, "title", "methodology.development.or.analysis", 
                "testing.macro.or.ecological.theory", "trends.over.time") 
names(qt_df)[2:4] <- c("methodology", "testing.theory", "temporal.trends")

qt_df <- gather(qt_df, key = "study_question", value = "v_2", 
                temporal.trends,testing.theory, methodology,  
         factor_key = T) %>%
  filter(v_2 == TRUE) %>%
  group_by(study_question) %>%
  summarise(count = n()) %>%
  mutate(percent = (count/sum(count)) * 100)

ggplot(data = qt_df, 
       aes(x = factor(study_question, 
                      levels = c("temporal.trends", "methodology",
                                 "testing.theory"), 
                      labels = c("temporal trends", "methodology",
                                 "testing theory")), 
           y = percent)) + 
  geom_bar(stat = "identity") + 
  scale_fill_brewer() + 
  xlab("Study Focus") + 
  ylab("Percent of Studies") + 
  theme_bw()


```



_________________________________________________________________________
### Old versions below here

## Types of questions  
Analyses may appear in this table or graph more than once.  For example, a study proposing a new method for estimating species richness and then using that method to test predictions about a productivity/richness relationship would appear in the counts in three columns.

```{r questions_by_data_type}
qt_df <- select(rev, "title", "year", "methodology.development.or.analysis", 
                "distribution", "abundance", "phenology", 
                "testing.macro.or.ecological.theory", "trends.over.time", 
                "data.type...what.where.when.only", 
                "data.type...detection...non.detection", 
                "data.type...abundance", "data.type...sampling.effort.reported", 
                "data.type...organized.data.collection.scheme") 
names(qt_df)[3:8] <- c("methodology", "distribution", "abundance",
                        "phenology", "testing.theory", "temporal.trends")

qt_df <- gather(qt_df, key = "study_question", value = "v_1", 
                methodology:temporal.trends, factor_key = T) %>%
  filter(v_1 == TRUE) %>%
  gather(key = "data.type", value = "v_2", 
         data.type...what.where.when.only:
           data.type...organized.data.collection.scheme, 
         factor_key = T) %>%
  filter(v_2 == TRUE)


kable(table(qt_df$data.type, qt_df$study_question), 
      caption = "The types of questions addressed by studies that use biological records, according to what type of data they use.")

### testing stuff
chisq.test(table(qt_df$data.type, qt_df$study_question), simulate.p.value = T)
```

```{r}
qt_df <- select(rev, "title", "year", "methodology.development.or.analysis", 
                "individual.species.analysis", "community.analysis", 
                "compiled.individual.species.analysis", "distribution", 
                "abundance", "phenology", "testing.macro.or.ecological.theory", 
                "trends.over.time") 
names(qt_df)[3:10] <- c("methodology", "individual.species", "community", 
                        "compiled.individual", "distribution", "abundance",
                        "phenology", "testing.theory")

qt_df <- gather(qt_df, key = "population", value = "v_1", individual.species, 
                compiled.individual, community, factor_key = T) %>%
  filter(v_1 == TRUE) %>%
  gather(key = "study_question", value = "v_2", distribution, abundance, 
         phenology, testing.theory, methodology, trends.over.time, 
         factor_key = T) %>%
  filter(v_2 == TRUE)


kable(table(qt_df$study_question, qt_df$population) / length(unique(qt_df$title)), digits = 2, 
      caption = "The types of questions addressed by studies that use biological records.  Values are the proportion of all studies.  Study populations are columns, topic of the study is in rows.")
```




## Taxonomic Group
```{r}
taxa <- select(rev, title, year, taxonomic.group, distribution, abundance, 
               phenology, testing.macro.or.ecological.theory, trends.over.time)

tax_names <- unlist(strsplit(taxa$taxonomic.group, "; "))
tax_names <- unique(tax_names)
tax_names <- tax_names[which(!is.na(tax_names))]
tax_names <- tax_names[order(tax_names)]

temp_df <- data.frame(matrix(nrow = nrow(taxa), ncol = length(tax_names)))
colnames(temp_df) <- tax_names
taxa <- cbind(taxa, temp_df)
rm(temp_df) 

for(i in 1:nrow(taxa)) {
  groups <- taxa$taxonomic.group[i]
  groups <- unlist(strsplit(groups, "; "))
  taxa[i, which(colnames(taxa) %in% groups)] <- T
}

taxa <- gather(taxa, key = "tax_group", value = "value", 
               all:wasps) %>%
  filter(value == T) 

table(taxa$tax_group)[order(table(taxa$tax_group), decreasing = T)]
```

```{r}
if(diag) {
  taxa <- gather(taxa, key = "question", value = "v_2", 
                 distribution:trends.over.time, factor_key = T) %>%
    filter(v_2 == T)
  
  table(taxa$tax_group, taxa$question)
}
```
Above is good.  Need to simplify taxonomic groups.  Might not end up using this table, b/c the patterns it shows probably have more to do with the type of data (abundance, semi-structure) than taxonomic group.  So, butterflies don't lend themselves to abundance and phenology studies, but the monitoring scheme used for butterflies does lend itself to those types of analyses.


--------------------------------------------------------------------

## Results format by data type
$H_01$: Different types of biological records data are not analyzed with different broad analysis approaches.

$H_a1$: Different types of biological records data are analyzed with different broad analysis approaches.

$H_02$: wh,wh,wh-only data is not analyzed with a different broad analysis approach than other data types.

$H_a2$: wh,wh,wh-only data is analyzed with inference and/or prediction less often than richer data are.  

$H_03$: The ?rate? at which wh,wh,wh-only data is analyzed with prediction is not greater than for other data types.

$H_a3$: The ?rate? at which wh,wh,wh-only data is analyzed with prediction is greater than for richer data types.

```{r analysis_approach_data_prep}
rf <- select(rev, c(title, data.type...what.where.when.only:
                      results.type...descriptive.only)) %>%
  gather(key = "data_type", value = "v_1", 
         data.type...what.where.when.only:data.type...detection...non.detection, 
         data.type...abundance:data.type...museum, 
         factor_key = T) %>%
  filter(v_1 ==T) %>%
  gather(key = "results_format", value = v_2, 
         results.type...inference:results.type...descriptive.only, 
         factor_key = T) %>%
  filter(v_2 == T)

# make names prettier
rf$results_format <- gsub("results.type...", "", rf$results_format)
rf$data_type <- gsub("data.type...", "", rf$data_type)

rf$results_format <- factor(as.character(rf$results_format), 
                            levels = c("inference", "prediction", "descriptive.only"), 
                            labels = c("inference", "prediction", "descriptive only"),
                            ordered = T)
```

```{r analysis_approach_by_data_type_assumptions_poisson_reg}
if(diag) {
  table(rf$data_type, rf$results_format)
  hist(table(rf$data_type, rf$results_format)) # doesn't look poisson-y
  
  addmargins(table(rf$data_type, rf$results_format))
  
  # check unconditional mean and variance - doesn't look the same
  mean(as.numeric(table(rf$data_type, rf$results_format)))
  var(as.numeric(table(rf$data_type, rf$results_format))) 

  
}
```

```{r resultsFormatByDataType}


if(diag) {
  pander(table(rev$results.type...descriptive.only), caption = "Descriptive Only Results (need at least 10 events/non evenst per predictor)")
}

dt_table <- table(rf$data_type)
names(dt_table) <- c("What, Where,\nWhen", "Abundance", "Known\nEffort", 
                     "Semi-\nStructured", "Visit-specific\nCovariates", 
                     "museum\nspecimen")
# pander(dt_table, caption = "Table of the gathered data_type column (compare to number of studies included)", keep.line.breaks = T)


kable(table(rf$data_type, rf$results_format), 
      caption = "Results formats produced using different types of data (this is counting studies, not counting analyses).  Is there a chi-squared-type test for when grouping variable categories are not mutually exclusive?")

```




Above looks ok, but probably need to do a chi-squared test or something to compare expected v. observed counts.  Or do glm with inference & hypothesis testing (or H testing, inference, and prediction) as positive class, “descriptive” as negative class, and data type as categorical predictors?  

### $H_02$

```{r results_by_data}
rev$quantitative <- rev$results.type...hypothesis.testing == T |
  rev$results.type...inference == T | rev$results.type...prediction == T

long_rev <- select(rev, title, quantitative, data.type...what.where.when.only:
                     data.type...detection...non.detection, 
                   data.type...sampling.effort.reported:data.type...museum) %>%
  gather(key = "data_type", value = "v_1", 
         data.type...what.where.when.only:data.type...museum) %>%
  filter(v_1 == T)

long_rev$quantitative <- gsub("TRUE", "inference/prediction", as.character(long_rev$quantitative))
long_rev$quantitative <- gsub("FALSE", "descriptive", as.character(long_rev$quantitative))
kable(table(long_rev$data_type, long_rev$quantitative))

```

```{r, echo = T}
if(diag) {
  table(rev$data.type...what.where.when.only)
  table(rev$data.type...abundance)
  table(rev$data.type...detection...non.detection)
  table(rev$data.type...sampling.effort.reported)
  table(rev$data.type...organized.data.collection.scheme)
  table(rev$data.type...visit.specific.covariates)
  table(rev$data.type...museum)
}
```

```{r fit_glm_analysis_approach}
results_data_model <- glm(quantitative ~ 
                            as.factor(data.type...what.where.when.only) + 
                            as.factor(data.type...abundance) + 
                            as.factor(data.type...detection...non.detection) + 
                            as.factor(data.type...sampling.effort.reported) + 
                            as.factor(data.type...organized.data.collection.scheme) + 
                            as.factor(data.type...visit.specific.covariates) + 
                            as.factor(data.type...museum), 
                          family = binomial, 
                          data = rev)
```

```{r glm_results}
# one_cov_mod <- glm(quantitative ~ data.type...what.where.when.only + 
#                             data.type...abundance, 
#                           family = binomial, 
#                           data = rev)

# summary(results_data_model)
# library(car)
# vif(results_data_model)
```











